---
title: "Machine Learning Assignment"
author: "Gregory Papaioannou"
date: "January 28, 2016"
output: html_document
---
#### Quantified Self Movement Data Analysis Report  
  
```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
library(parallel)
library(doParallel)
```
  
#### Parallel computing  
```{r}
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
```  
#### Load the data  
```{r}
train <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv')
test <- read.csv('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv')
```  
#### Get a first glimpse of the data  
```{r}  
# dim(trainRaw)
# dim(testRaw)
# summary(train)
# summary(test)
# sum(complete.cases(train))
```  
#### Start cleaning  
```{r}
# Drop the empty columns
train <- train[, colSums(is.na(train)) == 0] 
test <- test[, colSums(is.na(test)) == 0]

# Drop the columns that that do not contribute much to the accelerometer measurements
train <- train[, !grepl("^X|timestamp|window", names(train))]
test <- test[, !grepl("^X|timestamp|window", names(test))]

# Keep seperate the categorical variables 
train_user_name <- train$user_name
test_user_name <- test$user_name
classe <- train$classe

# Some more cleaning
train <- train[, sapply(train, is.numeric)]
test <- test[, sapply(test, is.numeric)]

train$user_name <- train_user_name
train$classe <- classe
test$user_name <- test_user_name
test <-subset(test, select=-problem_id)
```  
#### Cross validation sample  
```{r}
set.seed(13)
inTrain <- createDataPartition(train$classe, p=0.75, list=F)
trainData <- train[inTrain, ]
testData <- train[-inTrain, ]
```  
#### Create model with the Stochastic Gradient Boosting algorithm  
```{r}
# 5 folds
cv <- trainControl(method="cv", 5, allowParallel = TRUE)
model <- train(classe ~ ., data=trainData, method="rf", trControl=cv, ntree=250)
```  
#### Performance of the model on the validation data set  
```{r}
prediction <- predict(model, testData)
confusionMatrix(testData$classe, prediction)
```  
#### Accuracy  
```{r}
postResample(prediction, testData$classe)
```  
#### Estimated out-of-sample error  
```{r}
1 - as.numeric(confusionMatrix(testData$classe, prediction)$overall[1])
```  
#### Prediction for Test Data  
```{r}
predict(model, test)
```  
## Figures  
#### Correlation Matrix  
```{r}
correl <- cor(trainData[, sapply(train, is.numeric)])
corrplot(correl, method="color")
```  
#### Tree  
```{r}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel)
```  
  
